---
title: "Code blocks as examples"
format: html
editor: 
  markdown: 
    wrap: sentence
eval: false
echo: true
---

## Example 1

In the first example, we can imagine a scenario, where we are interested in spatiotemporal patterns of the *Picea* genus across North America for modern data and since the Last Glacial Maximum. Obtaining such data is straightforward:

```{r}
#| label: Example 1 - North America Picea
data_na_plots_picea <-
  # Access the VegVault
  vaultkeepr::open_vault(path = "<path_to_VegVault>") %>%
  # Start by adding dataset information
  vaultkeepr::get_datasets() %>%
  # Select both modern and paleo plot data
  vaultkeepr::select_dataset_by_type(
    sel_dataset_type = c(
      "vegetation_plot",
      "fossil_pollen_archive"
    )
  ) %>%
  # Limit data to North America
  vaultkeepr::select_dataset_by_geo(
    lat_lim = c(22, 60),
    long_lim = c(-135, -60)
  ) %>%
  # Add samples
  vaultkeepr::get_samples() %>%
  # Limit the samples by age
  vaultkeepr::select_samples_by_age(
    age_lim = c(0, 15e3)
  ) %>%
  # Add taxa & classify all data to a genus level
  vaultkeepr::get_taxa( classify_to = "genus") %>%
  # Extract only Picea data
  vaultkeepr::select_taxa_by_name(sel_taxa = "Picea") %>%
  vaultkeepr::extract_data()
```

## Example 2

In the second example, let's imagine we want to do Species Distribution Modeling for all plant taxa in the Czech Republic. We will extract modern plot-based data and Mean Annual temprature.

```{r}
#| label: Example 2 - CZ JSDM
data_cz_modern <-
  # Acess the VegVault file
  vaultkeepr::open_vault(path = "<path_to_VegVault>") %>%
  # Add the dataset information
  vaultkeepr::get_datasets() %>%
  # Select modern plot data and climate
  vaultkeepr::select_dataset_by_type(
    sel_dataset_type = c(
      "vegetation_plot",
      "gridpoints"
    )
  ) %>%
  # Limit data to Czech Republic
  vaultkeepr::select_dataset_by_geo(
    lat_lim = c(48.5, 51.1),
    long_lim = c(12, 18.9)
  ) %>%
  # Add samples
  vaultkeepr::get_samples() %>%
  # select only modern data
  vaultkeepr::select_samples_by_age(
    age_lim = c(0, 0)
  ) %>%
  # Add abiotic data
  vaultkeepr::get_abiotic_data() %>%
  # Select only Mean Anual Temperature (bio1)
  vaultkeepr::select_abiotic_var_by_name(sel_var_name = "bio1") %>%
  # add taxa
  vaultkeepr::get_taxa() %>%
  vaultkeepr::extract_data()
```

## Example 3

In the last example, let's imagine we want to reconstruct the Community Weighted Mean (CWM) of plant height for Latin America between 6-12 ka yr BP.

```{r}
#| label: Example 3 - LA CWM
data_la_traits <-
  # Acess the VegVault file
  vaultkeepr::open_vault(path = "<path_to_VegVault>") %>%
  # Add the dataset information
  vaultkeepr::get_datasets() %>%
  # Select modern plot data and climate
  vaultkeepr::select_dataset_by_type(
    sel_dataset_type = c(
      "fossil_pollen_archive",
      "traits"
    )
  ) %>%
  # Limit data to South and Central America
  vaultkeepr::select_dataset_by_geo(
    lat_lim = c(-53, 28),
    long_lim = c(-110, -38),
    sel_dataset_type = c(
      "fossil_pollen_archive",
      "traits"
    )
  ) %>%
  # Add samples
  vaultkeepr::get_samples() %>%
  # Limit to 6-12 ka yr BP
  vaultkeepr::select_samples_by_age(
    age_lim = c(6e3, 12e3)
  ) %>%
  # add taxa & clasify all data to a genus level
  vaultkeepr::get_taxa(classify_to = "genus") %>%
  # add trait information & clasify all data to a genus level
  vaultkeepr::get_traits(classify_to = "genus") %>%
  # Only select the plant height
  vaultkeepr::select_traits_by_domain_name(sel_domain = "Plant heigh") %>%
  vaultkeepr::extract_data()
```